---
title: "Homework 2"
author: "Meredith Tucker: mltucker2001"
date: "2/3/2020"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This homework is about syntax and basic plotting functions in R, and is **due by the end of Feb 7**. 

You can use the RMarkDown file with these homework instructions as a template for your work. 
Make sure to replace "Your Name" and "your_github_username" in the YAML with your name and Github username.

**Submission instructions**: Create a folder named `ds202_hw2`, and name the RMarkDown file including your solutions `hw2.Rmd` under this folder. For submission, create a GitHub repository named `ds202_hw2` under your account, and push both `hw2.Rmd` and the knitted `hw2.html` to GitHub before the deadline. I will assume you use the same GitHub username as for your HW1 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. Some questions will require the following answer format.

## Example question and answer
0\. What is the smallest positive odd integer? Storage it in a variable named `a0` and print it.

**Example answer**:
In the RMarkDown file, type in
<pre><code>```{r}
a0 <- 1
a0
```</code></pre>
The knitted html file will show:
```{r}
a0 <- 1
a0
```

## R commands
1. Today's temperature has a high of 24 and a low of 15 degree fahrenheit.  Find the corresponding degress in celsius and storage it in a (single) vector named `cel`. Print the variable.

```{r}
cel <- c(-4.4, -9.4)
cel
```
2. Create a vector of length 200. The odd entries should be 3 and even entries should be 1. Storage the created in a variable named `vec200` and print.

```{r}
vec200 <- rep(c(3,1), length.out=200)
vec200
```

3. How many perfect squares are there between 1 and 2001? Storage the answer in a  variable named `numPerfect` and print. You may want to look at `?round`.

```{r}
numPerfect <- floor(sqrt(2001))
numPerfect
```



## Cars and mileage
For this part we will use the `mpg` data set, you can access the data by first installing (using `install.packages('ggplot2')`) and then loading the package `ggplot2` (using `library(ggplot2)`) into your R session. The `mpg` data includes information about fuel economy of popular car models
for the years 1999 and 2008 as collected by the US Environmental Protection Agency. An answer should include a short paragraph of text for the explanation to a (human) reader, together with appropriate R code, outputs, and graphs.


```{r}
library(ggplot2)
str(mpg)
```

Most exercises are from the ggplot2 book section 2.2.1. You might find it helpful for working on the homework to read through this section. *However, all questions can be solved by using just the basic R graphics functions.*

1. What are the 3 most fuel-efficient models in terms of highway mileage? Storage the data frame (with 3 rows) containing all information of these three cars in a variable named `top3` and print.

```{r}

highway <- mpg$hwy
names(highway) <- mpg$model
top <- sort(highway, decreasing=T)[1:3]
top

top3 <- subset(mpg, hwy == 44 | hwy == 41)
top3

```

The models of cars with the top 3 hwy fuel efficency are shown above. They include a 1999 Jetta, a 1999 manual New Beetle, and a 1999 automatic New Beetle. The values of the top 3 fuel efficencies are 44, 44, and 41. 




2. How many different compact models are included in this dataset? Storage the answer in  a variable named `numCompact` and print.

```{r}

table(mpg$class== 'compact')

summary(mpg[mpg$class == 'compact', ])
numCompact <- 47
numCompact

```

There are 47 models of cars in the data set that have the class of "compact". The other class types are shown in the table above. There are 5 2seaters in the dataset. There are 41 midsize and 11 minivans. There are 33 pickups and 35 subcompacts. Finally, there are 62 SUVs.



3. Create a scatterplot of `hwy` and `cty` and describe the relationship. Why are there so few points visible? Explain your findings.

```{r}
plot( mpg$hwy, mpg$cty, main= "Relationship of hwy and cty", xlab= "hwy", ylab= "cty")

```

There appears to be a postive correlation between the fuel effiency in the city and on the highway. As the effiency on the highway increases so the the effiency in the city. The reason that there are so few data points in comparision to the number of entries in the data frame is probably the result of overlapping values on the scatter plot. If one vehicle gets a certain hwy and a certain cty, there is a large chance that another vehicle with the same hwy value will have the same cty value as the first vehicle. 


4. Are cars in 2008 more fuel efficient than cars in 1999? Use graphs *and* summary statistics to answer this question. Explain your findings.

```{r}
cars99 <- mpg[ mpg$year == 1999, ]
cars08 <- mpg[ mpg$year == 2008, ]

summary(cars99$hwy)
summary(cars99$cty)
summary(cars08$hwy)
summary(cars99$cty)

table1 <- c(mean(cars99$hwy), mean(cars99$cty), mean(cars08$hwy), mean(cars08$cty))

names(table1) <- c( "1999 Highway", "1999 City", "2008 Highway", "2008 City")
table1


barplot(table1, main= "Fuel Efficiency")


boxplot(cars99$hwy, cars99$cty, cars08$hwy, cars08$cty, main= 'Fuel Effieciency', names= c('1999 Highway', '1999 City', '2008 Highway', '2008 City'), col = c('red', 'blue'))


```

The cars in 2008 had a higher average highway efficiency than the highway efficiency in 1999. The cars in 1999 had a better efficiency in the city then the cars in 2008 did in the city. This can be seen in both the table and the box plot.


Note: your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission. Make sure that you don't use any file structure that depends on your computing environment.